Labeled object representation and manipulation with ITK

### 2007-08-30 article/Article.tex
some corrections on the computation details
darcs-hash:20070830135535-2fc9d-98127f46b4d9fb87085437cec0d521419f0906a1.gz

@@

\begin{eqnarray}
V_n(r) & = {\pi^\frac{n}{2}r^n\over\Gamma(\frac{n}{2} + 1)}
\end{eqnarray}
where $V_n$ is the volume of the hypersphere $n$ is the image dimension,
and $r$ is the radius of the hypersphere.

@@

$n!!$ is the double factorial, defined as

@@

\begin{eqnarray}
A_n(r) & = {\frac{n V_n}{r}}
\end{eqnarray}

Where $A_n$ is the area of the hypersphere.

\begin{eqnarray}
R & = {\frac{a}{A_n(r)}}
\end{eqnarray}

where $R$ is the roundness, $a$ is the measured area of the object, and the $r$ is the radius of an hypersphere with the same volume than the object.

>>>>

Central image moments for grayscale images are usually computed as

\begin{eqnarray}
Cm_{i,j} & = & \frac{S_{i,j}}{M} - Cg_i.Cg_j \\
S_{i,j}  & = & \sum_{p \in D} \left( I(p).p_i.p_j \right)
\end{eqnarray}

where $S_{i,j}$ is the central moment, $D$ is the domain of definition of the
image $I$, $I(p)$ is the pixel value of the image at the position $p$,
$0 \leq i < n$, $0 \leq i < n$, $n$ is the image dimension, $p_i$ is the
physical position on the axis $i$, $M$ is the total mass, $Cg$ is the center of
gravity. With binary images, $I(p)$ is either $0$ 
if $p$ is outside the object, or $1$ if $p$ is inside.

The complexity is $O(N_{P_I})$, where $N_{P_I}$ is the number of pixels in the image.

With the run-length encoding of the binary objects, the complexity can be decreased
to $O(N_{L_O})$, where $N_{L_O}$ is the number of lines in the object.

\begin{eqnarray}
S_{i,j} & = & \sum_{p \in D} \left( I(p).p_i.p_j \right) \\ \nonumber
         & = & \sum_{p \in O} \left( p_i.p_j \right) \\ \nonumber
         & = & \sum_{L \in O}  \sum_{p \in L} \left( p_i.p_j \right)
\end{eqnarray}

Where $O$ is a binary object in the image, and $L$ is a line of the object $O$, encoded
with the run-length encoding. In a line, $p_i$ is a constant if $i > 0$, so, $S_{i,j}$
can be written:

\begin{eqnarray}
S_{i,j} = \begin{cases}
  \underset{L \in O}\sum \left( l_L.p_i.p_j \right)                & \text{if $i > 0$ and $j > 0$}, \\
  \underset{L \in O}\sum \left( l_L.p_i.\underset{p \in L}\sum p_0 \right)  & \text{if $i > 0$ and $j = 0$}, \\
  \underset{L \in O}\sum \left( l_L.p_j.\underset{p \in L}\sum p_0 \right)  & \text{if $i = 0$ and $j > 0$}, \\
  \underset{L \in O}\sum \underset{p \in L}\sum p_0^2            & \text{if $i = j = 0$}.
           \end{cases}
\end{eqnarray}

It is known that
\begin{eqnarray}
\label{equ:sumx}
\sum_{x=0}^n x   & = & \frac{n(n+1)}{2} \\
\label{equ:sumx2}
\sum_{x=0}^n x^2 & = & \frac{n(n+1)(2n+1)}{6}
\end{eqnarray}

In order to use those formulae in the computation of $S_{i,j}$, the physical position has to be expanded in:
\begin{eqnarray}
\label{equ:expandedPosition}
p_j = o_j + s_j i_j
\end{eqnarray}
where $o_j$ is the origin of the line on the axis $j$, $s_j$ is the spacing on the axis $j$, and
$i_j$ is the index on the axis $j$.

With equations \ref{equ:expandedPosition}, \ref{equ:sumx} and \ref{equ:sumx2}, it is easy to remove the loop in the computation of the sum of physical positions of the axis $0$:

\begin{eqnarray}
\sum_{p \in L} p_0  & = & \sum_{i_0=0}^{l_L-1} (o_0 + s_0 i_0) \\ \nonumber
                    & = & l_L.o_0 + s_0\sum_{i_0=0}^{l_L-1} i_0 \\ \nonumber
                    & = & l_L.o_0 + s_0\left( \frac{(l_L-1)l_L}{2} \right) \\ \nonumber
                    & = & l_L \left( o_0 + \frac{s_0(l_L-1)}{2} \right)
\end{eqnarray}

where $l_L$ is the length of the line (in pixels), and in the sum of the square of the physical positions of the axis $0$:

\begin{eqnarray}
\sum_{p \in L} p_0^2  & = & \sum_{i_0=0}^{l_L-1} ( o_0 + s_0 i_0 )^2 \\ \nonumber
                      & = & \sum_{i_0=0}^{l_L-1} ( o_0^2 + s_0^2 i_0^2 + 2 o_0 i_0) \\ \nonumber
                      & = & l_L.o_0^2 + s_0^2\sum_{i_0=0}^{l_L-1} i_0^2 + 2 o_0 \sum_{i_0=0}^{l_L-1} i_0 \\ \nonumber
                      & = & l_L.o_0^2 + s_0^2 \left( \frac{(l_L-1)l_L(2l_L-1)}{6} \right) + 2 o_0 \left( \frac{(l_L-1)l_L}{2} \right) \\ \nonumber
                      & = & l_L\left( o_0^2 + (l_L-1)\left(\frac{s_0^2(2l_L-1)}{6}  + o_0 \right) \right)
\end{eqnarray}

Finally, $S_{i,j}$, used in the computation of the central moments, is computed as:
\begin{eqnarray}
S_{i,j} = \begin{cases}
  \underset{L \in O}\sum \left( l_L.p_i.p_j \right)                & \text{if $i > 0$ and $j > 0$}, \\
  \underset{L \in O}\sum \left( p_i.l_L^2 \left( o_0 + \frac{s_0(l_L-1)}{2} \right) \right) & \text{if $i > 0$ and $j = 0$}, \\
  \underset{L \in O}\sum \left( p_j.l_L^2 \left( o_0 + \frac{s_0(l_L-1)}{2} \right) \right) & \text{if $i = 0$ and $j > 0$}, \\
  \underset{L \in O}\sum \left( l_L\left( o_0^2 + (l_L-1)\left(\frac{s_0^2(2l_L-1)}{6}  + o_0 \right) \right) \right)            & \text{if $i = j = 0$}.
           \end{cases}
\end{eqnarray}

@@

\begin{eqnarray}
\label{equ:hypersphereVolume}
V_n(r) = \frac{\pi^\frac{n}{2}r^n}{\Gamma(\frac{n}{2} + 1)}
\end{eqnarray}
where $V_n$ is the volume of the hypersphere $n$ is the image dimension,
and $r$ is the radius of the hypersphere.

@@

$n!!$ is the {\em double factorial}, defined as:

@@

\begin{eqnarray}
A_n(r) = {\frac{n V_n}{r}}
\end{eqnarray}

where $A_n$ is the area of the hypersphere.

\begin{eqnarray}
R = {\frac{a}{A_n(r)}}
\end{eqnarray}

where $R$ is the roundness, $a$ is the measured area of the 
object\footnote{More details about perimeter estimation will be published 
in another article.}, and the $r$ is the radius of an hypersphere with the
same volume than the object, computed using equation
\ref{equ:hypersphereVolume}.
